---
title: "Draft: Results"
format: 
  html:
    code-fold: true
    code-tools: true
---

Using the bridge inspection data from the National Bridge Inventory (NBI), an analysis of all bridges within the Omaha area can be conducted. This is done by identifying bridges with the Douglas County area code (055) and surrounding areas that serve Omaha traffic. Key variables from the NBI include deck condition, superstructure and substructure condition rating, structural evaluation, age of structure, average daily traffic, and bridge condition will be used in the primary analysis.

```{r}
#| include: FALSE
library(tidyverse)
library(ggplot2)
nbi_data <- read.csv("NE24.txt", header = TRUE)
data <- nbi_data %>%
  filter(COUNTY_CODE_003 == "55")
```

```{r}
ggplot(data, aes(x = LOWEST_RATING)) +
  geom_histogram(binwidth = 1, fill = "grey", color = "black") +
  stat_bin(
    binwidth = 1, 
    geom = "text", 
    aes(label = after_stat(count)), 
    vjust = -0.5, 
    size = 3.5
  ) +
  scale_x_continuous(
    breaks = seq(
      from = floor(min(nbi_data$LOWEST_RATING, na.rm = TRUE)),
      to = ceiling(max(nbi_data$LOWEST_RATING, na.rm = TRUE)),
      by = 1
    )
  ) +
  labs(
    title = "Distribution of Bridge Condition Ratings",
    x = "Lowest Condition Rating",
    y = "Count"
  ) +
  theme_minimal()
```

It's found that there are 15 bridges (2.9%) in Douglas County that have at least one condition rating of 4, indicating "poor condition" where there is advanced section loss, deterioration, spalling, or scour. There are 0 bridges with a rating worse than 4, meaning that no bridges are considered to be in "imminent" failure condition or already failed/collapsed. Overall, bridges that are considered to be in poor condition are more often older bridges, with the average year built being 1942. The average year built for bridges in fair or good conditions is 1980, where bridges with fair or good conditions were more likely to be built recently.

```{r}
ggplot(data, aes(x = YEAR_BUILT_027, y = BRIDGE_CONDITION, fill = BRIDGE_CONDITION)) +
  geom_boxplot() +
  labs(
    title = "Bridge Condition by Construction Year",
    x = "Year Built",
    y = "Bridge Condition Rating",
    fill = "Condition"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray40"),
    legend.position = "right",
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  scale_x_continuous(
    breaks = seq(
      floor(min(nbi_data$YEAR_BUILT_027, na.rm = TRUE)/10)*10,
      ceiling(max(nbi_data$YEAR_BUILT_027, na.rm = TRUE)/10)*10,
      10
  )
)
a <- data %>% 
  filter(LOWEST_RATING <= 4) %>% 
  summarise(Avg_Year_Poor = round(mean(YEAR_BUILT_027, na.rm = TRUE), 0)) %>%
  pull(Avg_Year_Poor)

b <- data %>% 
  filter(LOWEST_RATING > 4) %>% 
  summarise(Avg_Year_Not_Poor = round(mean(YEAR_BUILT_027, na.rm = TRUE), 0)) %>% 
  pull(Avg_Year_Not_Poor)

data %>%
  group_by(LOWEST_RATING) %>%
  summarise(mean_ADT = sum(ADT_029, na.rm = TRUE)) %>%
  ggplot(aes(x = LOWEST_RATING, y = mean_ADT)) +
  geom_col(fill = "grey") +
  geom_text(aes(label = round(mean_ADT)), vjust = -0.5) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  labs(
    title = "Average Traffic by Bridge Condition Rating",
    x = "Lowest Condition Rating",
    y = "Mean Daily Traffic"
  ) +
  theme_minimal()
```

For the bridges that are in poor condition, there is very low average daily traffic compared to bridges with higher condition ratings.